Extreme Value Chapter

### 2022-03-01 Ch3_samplesize/Figures_Ch3.Rmd
Working slowly on the details

---
title: "Extreme Value Chapter"
author: "Javiera Rudolph"
output:
  html_document:
    toc: yes
---

@@

```{r libraries}
library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(purrr)
library(MetBrewer)
library(stringr)
```

@@

pal_sampsize <- met.brewer("Hokusai3", n = 6)
pal_threshold <- met.brewer("Derain", n = 7)
pal_mixture <- met.brewer("Hokusai1", n = 4)
theta_color <- "steelblue"

@@

# The color associated to each mixture component
dens_cols <- c("#264653", "#2a9d8f", "#f4a261", "#e76f51")

@@

```{r fig.width=6, fig.height=2}
thetas %>% 
  ggplot(., aes(x = thresh_tests, y = theta)) +
  geom_point(size = 4, shape = 18, color = theta_color) +
  labs(x = "Long-distance level", y = "Theta") +
  theme_bw() -> a

thetas %>% 
  ggplot(., aes(x = thresh_tests, y = theta)) +
  geom_point(size = 4, shape = 18, color = theta_color) +
  labs(x = "Long-distance level", y = "Theta", Title = "Log Scale") +
  theme_bw() +
  scale_y_log10() -> b

plot_grid(a, b, ncol = 2)

@@

Now, we explore how the estimates compare. These are already in log scale because they are small:

```{r fig.width=6, fig.height=4}
nreps_mles_df %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_point(alpha = 0.8, size = 3, color = "grey27") +
  labs(x = "Long-distance level", y = "Estimated thetas", title = "Simple Lomax") +
  scale_x_continuous(breaks = thresh_tests) +
  theme_bw() -> a

@@

Since there are so many points, we can see them a little more scattered with jitter:

```{r fig.width=6, fig.height=4}
nreps_mles_df %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_jitter(alpha = 0.8, size = 3, color = "grey27") +
  labs(x = "Long-distance level", y = "Estimated thetas", title = "Simple Lomax") +
  scale_x_continuous(breaks = thresh_tests) +
  theme_bw() -> a

@@

But we are seeing aggregated data, and we tested at different sample sizes:

```{r fig.width=8, fig.height=4}
nreps_mles_df %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_jitter(alpha = 0.8, size = 3, aes(color = factor(samp_n_tests))) +
  scale_color_manual(values = pal_sampsize) +
  labs(x = "Long-distance level", y = "Estimated thetas", title = "Simple Lomax") +
  scale_x_continuous(breaks = thresh_tests) +
  theme_bw() +
  theme(legend.position = "none") -> a

@@

Can I see this as boxplots or something else? To see the differences in sample size more clearly.

```{r fig.width=8, fig.height=4}
nreps_mles_df %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat, fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  scale_fill_manual(values = pal_sampsize) +
  scale_color_manual(values = pal_sampsize) +
  labs(x = "Long-distance level", y = "Theta hat", title = "Simple Lomax") +
  theme_bw() +
  theme(legend.position = "none") -> a

@@

```{r eval = FALSE, fig.width=8, fig.height=4}

# See it compared to the original theta
nreps_mles_df %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat - log(theta), fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, color = "red") +
  scale_fill_manual(values = pal_sampsize) +
  scale_color_manual(values = pal_sampsize) +
  labs(x = "Long-distance level", y = "log(theta_hat) - log(theta)", title = "Simple Lomax") +
  theme_bw() +
  theme(legend.position = "none") -> a

@@

@@

>>>>

---
title: "Extreme Value Chapter"
author: "Javiera Rudolph"
output:
  html_document:
    toc: yes
    self_contained: false
---

@@

```{r libraries}
library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(purrr)
library(MetBrewer)
library(stringr)
library(latex2exp)
```

@@

pal_sampsize <- met.brewer("Hokusai3", n = 6)
pal_threshold <- met.brewer("Derain", n = 7)
pal_mixture <- met.brewer("Hokusai1", n = 4)
theta_color <- "steelblue"
pal_bias <- met.brewer("Lakota", n=5)

@@

# The color associated to each mixture component
dens_cols <- c("#264653", "#2a9d8f", "#f4a261", "#e76f51")
# dens_cols <- pal_mixture

@@

```{r fig.width=5, fig.height=2.5}
# thetas %>% 
#   ggplot(., aes(x = thresh_tests, y = theta)) +
#   geom_point(size = 4, shape = 18, color = theta_color) +
#   labs(x = "Long-distance level", y = "Theta") +
#   theme_bw() -> a

thetas %>% 
  ggplot(., aes(x = thresh_tests, y = log(theta))) +
  geom_point(size = 3, shape = 18, color = theta_color) +
  labs(x = "Distance level", y = "log(theta)") +
  scale_x_continuous(breaks = thresh_tests) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 30)) -> b
b
# plot_grid(a, b, ncol = 2)

@@

```{r eval = FALSE, fig.width=6, fig.height=4}
nreps_mles_df %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_point(alpha = 0.8, size = 3, color = "grey27") +
  labs(x = "Long-distance level", y = "Estimated thetas", title = "Simple Lomax") +
  scale_x_continuous(breaks = thresh_tests) +
  theme_bw() -> a

@@

```{r eval = FALSE, fig.width=6, fig.height=4}
nreps_mles_df %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_jitter(alpha = 0.8, size = 3, color = "grey27") +
  labs(x = "Long-distance level", y = "Estimated thetas", title = "Simple Lomax") +
  scale_x_continuous(breaks = thresh_tests) +
  theme_bw() -> a

@@

```{r eval = FALSE, fig.width=8, fig.height=4}
nreps_mles_df %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_jitter(alpha = 0.8, size = 3, aes(color = factor(samp_n_tests))) +
  scale_color_manual(values = pal_sampsize) +
  labs(x = "Long-distance level", y = "Estimated thetas", title = "Simple Lomax") +
  scale_x_continuous(breaks = thresh_tests) +
  theme_bw() +
  theme(legend.position = "none") -> a

@@

```{r eval = FALSE, fig.width=8, fig.height=4}
nreps_mles_df %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat, fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  scale_fill_manual(values = pal_sampsize) +
  scale_color_manual(values = pal_sampsize) +
  labs(x = "Long-distance level", y = "Theta hat", title = "Simple Lomax") +
  theme_bw() +
  theme(legend.position = "none") -> a

@@

```{r fig.width=8, fig.height=4}

# See it compared to the original theta
nreps_mles_df %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat - log(theta), fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, color = "red") +
  scale_fill_manual(values = pal_sampsize) +
  scale_color_manual(values = pal_sampsize) +
  labs(x = "Distance level", y = TeX(r'($log(\hat{\theta} / \theta)$)'), subtitle = "Simple Lomax") +
  theme_bw() +
  theme(legend.position = "none") -> a

nreps_mles_df %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat2 - log(theta), fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, color = "red") +
  scale_fill_manual(values = pal_sampsize, name = "Sample Size") +
  scale_color_manual(values = pal_sampsize, name = "Sample Size") +
  labs(x = "Distance level", y = TeX(r'($log(\hat{\theta} / \theta)$)'), subtitle = "GLM Lomax") +
  theme_bw() +
  theme(legend.position = "right") -> b

@@

```{r fig.width=10, fig.height=5}

# See it compared to the original theta
nreps_mles_df %>% 
  filter(., thresh_tests < 250) %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat - log(theta), fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, color = "red") +
  scale_fill_manual(values = pal_sampsize) +
  scale_color_manual(values = pal_sampsize) +
  labs(x = "Distance level", y = TeX(r'($log(\hat{\theta} / \theta)$)'), subtitle = "Simple Lomax") +
  theme_bw() +
  theme(legend.position = "none") -> a1

nreps_mles_df %>% 
  filter(., thresh_tests >= 250) %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat - log(theta), fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, color = "red") +
  scale_fill_manual(values = pal_sampsize) +
  scale_color_manual(values = pal_sampsize) +
  labs(x = "Distance level", y = " ", subtitle = " ") +
  theme_bw() +
  theme(legend.position = "none") -> a2

a <- plot_grid(a1, a2, ncol = 2)

nreps_mles_df %>% 
  filter(., thresh_tests < 250) %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat2 - log(theta), fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, color = "red") +
  scale_fill_manual(values = pal_sampsize, name = "Sample Size") +
  scale_color_manual(values = pal_sampsize, name = "Sample Size") +
  labs(x = "Distance level", y = TeX(r'($log(\hat{\theta} / \theta)$)'), subtitle = "GLM Lomax") +
  theme_bw() +
  theme(legend.position = "none") -> b1

nreps_mles_df %>% 
  filter(., thresh_tests >= 250) %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat2 - log(theta), fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, color = "red") +
  scale_fill_manual(values = pal_sampsize, name = "Sample Size") +
  scale_color_manual(values = pal_sampsize, name = "Sample Size") +
  labs(x = "Distance level", y = " ", subtitle = " ") +
  theme_bw() +
  theme(legend.position = "none") -> b2

b2

b <- plot_grid(b1, b2, ncol = 2)

ableg <- get_legend(b + theme(legend.position = "right"))

abplot <- plot_grid(a, b, nrow = 2, labels = c("A", "B"))

plot_grid(abplot, ableg, ncol = 2, rel_widths = c(1, 0.2))

@@

```{r eval = FALSE, fig.width=10}
corr_b %>% 
  bind_rows(., corr_a) %>% 
  right_join(., thetas) %>%
  group_by(thresh_tests, samp_n_tests, type, correctiontype) %>% 
  summarize(Eval_theta_hat = mean(value),
            theta = mean(theta),
            bias = Eval_theta_hat - theta) %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  mutate(labelsamp = paste("n =", samp_n_tests),
         labelsamp = factor(labelsamp, levels = c("n = 80", "n = 200", "n = 500", "n = 800", "n = 1000", "n = 1600"))) %>% 
  mutate(bootcat = case_when(type == "Corrected" & correctiontype == "Nonparametric" ~ "NP - C",
                             type == "Corrected" & correctiontype == "Parametric" ~ "P - C",
                             type == "Original" & correctiontype == "Nonparametric" ~ "NP - UC",
                             type == "Original" & correctiontype == "Parametric" ~ "P - UC")) %>% 
  ggplot(., aes(y = bias, x = thresh_tests)) +
  facet_wrap(~labelsamp) +
  geom_jitter(aes(color = bootcat), size = 4, alpha = 0.8, width = 0.1) +
  geom_hline(yintercept = 0, color = "red") +
  scale_color_manual(values = pal_bias, name = "") +
  labs(x = "Long-distance level", y = "Bias") +
  theme_bw()
```

# Figure 6
Compare to the GP

```{r fig.width=8, fig.height=4}

GP_mles %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log(quant_theta/theta), fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, color = "red") +
  scale_fill_manual(values = pal_sampsize, name = "Sample Size") +
  scale_color_manual(values = pal_sampsize, name = "Sample Size") +
  labs(x = "Long-distance level", y = "log(GP_theta/theta)", title = "Generalized Pareto") +
  theme_bw() +
  theme(legend.position = "none") -> a

```{r fig.width=8, fig.height=4}

nreps_mles_df %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log_St_hat2 - log(theta), fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, color = "red") +
  scale_fill_manual(values = pal_sampsize, name = "Sample Size") +
  scale_color_manual(values = pal_sampsize, name = "Sample Size") +
  labs(x = "Long-distance level", y = "log(theta_hat) - log(theta)", title = "GLM Lomax") +
  theme_bw() +
  theme(legend.position = "right") +
  lims(y = c(-5, 5)) -> b

GP_mles %>% 
  mutate(samp_n_tests = factor(samp_n_tests),
         thresh_tests = factor(thresh_tests)) %>% 
  ggplot(., aes(x = thresh_tests, y = log(quant_theta/theta), fill = samp_n_tests, color = samp_n_tests)) +
  # facet_wrap(~samp_n_tests, nrow = 1) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, color = "red") +
  scale_fill_manual(values = pal_sampsize, name = "Sample Size") +
  scale_color_manual(values = pal_sampsize, name = "Sample Size") +
  labs(x = "Long-distance level", y = "log(GP_theta/theta)", title = "Generalized Pareto") +
  theme_bw() +
  theme(legend.position = "none")+
  lims(y = c(-5, 5)) -> a
